Free surface computations in mantle convection models

### 2015-09-15 free-surface-paper.tex
Address comments

Finally, we discuss the implementation of a free surface in the open source, community based
mantle convection software \texttt{ASPECT}.
\end{abstract}

@@

Finally, one can use a true free surface, where a stress-free boundary condition is applied on the boundary 
of the domain. In this case, there can be flow in and out of the domain, so the boundary must move in time.
A true free surface has mathematical elegance in that the boundary condition of the domain more closely 
matches the boundary conditions which one is trying to model, but it typically requires a deformable 
domain with frequent remeshing to avoid ill-conditioned cells.

@@

Finally, we describe the implementation of a free surface in the mantle convection software \texttt{ASPECT}.

@@

\section{Implementation in \texttt{ASPECT} }
\label{sec:implementation}
We have implemented the ability to run free surface simulations in the mantle convection software \texttt{ASPECT} \citep{kronbichler2012high}. 
\texttt{ASPECT} is designed to be highly flexible and modular, with the ability for user-defined rheologies, geometries, and gravity models. 
The implementation of the free surface, therefore, needs to be general enough to work for many combinations of these models, including those 
which may not have been written yet. In particular, it cannot rely on assumptions regarding the shape of the domain.

Furthermore, \texttt{ASPECT} allows for computations in both 2D and 3D, so the implementation must be sufficiently dimension
independent to work in both cases. We implement the free surface within an arbitrary Lagrangian-Eulerian (ALE) formulation
\citep[e.g][]{fullsack1995arbitrary,donea2004encyclopedia}.

\texttt{ASPECT} is a parallelized, distributed memory code with adaptive mesh refinement.
The free surface implementation works with these features.  We store the mesh vertex positions 
in a fully distributed vector. This vector is continually updated and redistributed across 
processes upon mesh adaptation (which is handled by the adaptive octree library \texttt{p4est}).
We also provide adaptive refinement indicators based on being near to the free surface or 
when the free surface slope is steep to allow for accurate interface tracking.

@@

Finally, we have described the implementation of free surface boundary conditions in 
the open source mantle convection software \texttt{ASPECT}. Both the quasi-implicit 
scheme and the nonstandard finite difference scheme are available. The implementation is 
sufficiently general to accommodate many different geometies, rheologies, and 
gravity models. Furthermore, it runs in parallel and with adaptive mesh refinement.

>>>>

Finally, we discuss the implementation of a free surface in the open source, community based
mantle convection software \texttt{Aspect}.
\end{abstract}

@@

Finally, one can use a true free surface, where a stress-free boundary condition is applied on the boundary 
of the domain. In this case, there can be flow in an out of the domain, so the boundary must move in time.
A true free surface has mathematical elegance in that the boundary condition of the domain more closely 
matches the boundary conditions which one is trying to model, but it typically requires a deformable 
domain with frequent remeshing to avoid ill-conditioned cells.

@@

Finally, we describe the implementation of a free surface in the mantle convection software \texttt{Aspect}.

@@

\section{Implementation in \texttt{Aspect} }
\label{sec:implementation}
We have implemented the ability to run free surface simulations in the mantle convection software \texttt{Aspect} \citep{kronbichler2012high}. 
\texttt{Aspect} is designed to be highly flexible and modular, with the ability for user-defined rheologies, geometries, and gravity models. 
The implementation of the free surface, therefore, needs to be general enough to work for many combinations of these models, including those 
which may not have been written yet. In particular, it cannot rely on assumptions regarding the shape of the domain.

Furthermore, \texttt{Aspect} allows for computations in both 2D and 3D, so the implementation must be sufficiently dimension
independent to work in both cases. We implement the free surface within an arbitrary Lagrangian-Eulerian (ALE) formulation
\citep[e.g][]{fullsack1995arbitrary,donea2004encyclopedia}.

\texttt{Aspect} is a parallelized, distributed memory code with adaptive mesh refinement.
The free surface implementation works with these features.  We store the mesh vertex positions 
in a fully distributed vector. This vector is continually updated and redistributed across 
processes upon mesh adaptation (which is handled by the adaptive octree library \texttt{p4est}).
We also provide adaptive refinement indicators based on being near to the free surface or 
when the free surface slope is steep to allow for accurate interface tracking.

@@

Finally, we have described the implementation of free surface boundary conditions in 
the open source mantle convection software \texttt{Aspect}. Both the quasi-implicit 
scheme and the nonstandard finite difference scheme are available. The implementation is 
sufficiently general to accommodate many different geometies, rheologies, and 
gravity models. Furthermore, it runs in parallel and with adaptive mesh refinement.
